Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-02-26 automatica/paper.tex
sections 2 and 3

@@

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable~\cite{Astrom08,
astrom1997computer}:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad 
\vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q,\,
\mat{A} \in \mathbb{R}^{p \times p},\,
\mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

Introduce set $X=\{ \vec{x} : \mat{C}\vec{x} \leq \vec{d}\}$. 
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into:
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', \text{ such that } \bigcup_{k \in [0\ \ \infty)}\mat{J}^k \subseteq \mathcal{J}, 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in Section \ref{sec:eigensolver} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}). 
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}).
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the sound over-approximation of the double description algorithm \cite{fukuda1996double} from Section \ref{sec:dd} (line \ref{alg:getVertices}). It is worth noting that this algorithm uses a simplex to seed it: we will first discuss the simplex algorithm, which is also needed at the next step, and then the vertex enumeration algorithm used at this stage. 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in Section \ref{sec:simplex}.  The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j : \vec{v}_i \in V_0 \wedge \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}

@@

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$, we must be able to ensure that
certain properties of the plant hold at any given point in time.  In order
to do so, we accelerate the dynamics of the plant according to the model
semantics in Sec.~\ref{sec:model} and compute the state
$\vec{x}_T$ at a given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

@@

The final case we analyze is that of hybrid-time models, where the plant
dynamics evolve in continuous time while the feedback dynamics evolve in
discrete time.  A~new input value is produced only at every sample time but
implemented over the continuous-time plant dynamics.
%
\begin{theorem}
The expression
%
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
%
with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in
\eqref{eq:continuous_tube_param}, and where $\mat{A}_d=\mat{A}_{T_s}$,
$\mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with
sample time $T_s$, is the state of the model
%
\begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s \,. 
\end{align*}
%
\end{theorem}
\begin{proof}
%
Let $\mat{K}$ be a feedback controller such that
$\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$.  Let the output of the
feedback system be a zero order holder such that
$\vec{u}(t'+kT_s)=\vec{u}(kT_s), \text{ for } 0 \leq t' < T_s$.  The closed
loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%
\begin{align*}
\vec{x}_{k,t'}=\mat{A}_{t'}\vec{x}_k-\mat{B}_{t'}\mat{K}\vec{x}_k = (\mat{A}_{t'}-\mat{B}_{t'}\mat{K})\vec{x}_k,
\end{align*}
where $0 \leq t' < T_s$
We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
\begin{align*}
\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
\end{align*}
which accelerating results and setting $T=kT_s+t'$ becomes
\begin{align}
\label{eq:feedback_sampled_cont}
\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
\vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
\label{eq:feedback_cont}
\end{align}
\end{proof}

>>>>

\renewcommand{\note}[1]{\textcolor{red}{[#1]}}

@@

\note{Add note explaining relationship of this work with previously published papers. }

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable~\cite{Astrom08,
astrom1997computer}: \note{Pls eliminate vector quantities, unless strictly needed.}
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad 
\vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q,\,
\mat{A} \in \mathbb{R}^{p \times p},\,
\mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs
\note{Do we need the output map? Where is it used?}
\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  
\note{Next needed?}
In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

Introduce set $X=\{ \vec{x} : \mat{C}\vec{x} \leq \vec{d}\}$. \note{This notation clashes with that of the output map for the LTI model. }
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
\note{The second condition has already been defined above.}
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', \text{ such that } \bigcup_{k \in [0\ \ \infty)}\mat{J}^k \subseteq \mathcal{J}, 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in Section \ref{sec:eigensolver} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}). \note{wrong ref.}
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the sound over-approximation of the double description algorithm \cite{fukuda1996double} from Section \ref{sec:dd} (line \ref{alg:getVertices}). 
\note{wrong refs.} 
It is worth noting that this algorithm uses a simplex to seed it: we will first discuss the simplex algorithm, which is also needed at the next step, and then the vertex enumeration algorithm used at this stage. 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in Section \ref{sec:simplex}.  \note{fix ref.} The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j : \vec{v}_i \in V_0 \wedge \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}

@@

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  
\note{Unlike the CAV18 paper, these gain matrices have not been discussed yet. }
However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$ \note{Do we need this second matrix?}, we must be able to ensure that
certain properties of the plant hold at any given point in time.  In order
to do so, we accelerate the dynamics of the plant according to the model
semantics in Sec.~\ref{sec:model} and compute the state
$\vec{x}_T$ at a given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

@@

The final case we analyze is that of hybrid-time models, where the plant
dynamics evolve in continuous time while the feedback dynamics evolve in
discrete time.  A~new input value is produced only at every sample time but
implemented over the continuous-time plant dynamics.
%
\note{Matrix K has not been introduced yet. }
\begin{theorem}
The expression
%
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
%
with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in
\eqref{eq:continuous_tube_param}, and where $\mat{A}_d=\mat{A}_{T_s}$,
$\mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with
sample time $T_s$, is the state of the model
%
\begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s \,. 
\end{align*}
%
\end{theorem}
\begin{proof}
%
Let $\mat{K}$ be a feedback controller such that
$\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$.  Let the output of the
feedback system be a zero order holder such that
$\vec{u}(t'+kT_s)=\vec{u}(kT_s), \text{ for } 0 \leq t' < T_s$.  The closed
loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%
\begin{align*}
\vec{x}_{k,t'}=\mat{A}_{t'}\vec{x}_k-\mat{B}_{t'}\mat{K}\vec{x}_k = (\mat{A}_{t'}-\mat{B}_{t'}\mat{K})\vec{x}_k,
\end{align*}
where $0 \leq t' < T_s$
We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
\begin{align*}
\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
\end{align*}
which accelerating results and setting $T=kT_s+t'$ becomes
\begin{align}
\label{eq:feedback_sampled_cont}
\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
\vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
\label{eq:feedback_cont}
\end{align}
\end{proof}
